---
title: "wp_spectral_analysis"
author: "Charles Rivera"
date: "8/7/23"
output: html_notebook
---

Load dependencies
```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(tidyr)
library(dplyr) 
```

Functions to read in csv files from directory
```{r}
spectral_data = read.csv('/Users/charles/Desktop/wp_asd_spectra/wp_summer_2021/wp_summer.csv', sep=',', row.names=1)
```


Remove white references
```{r}
#===========
# Specified string to exclude
specifiedString <- "WR"
#
# # Get the row indices that match the specified string
matchingRows <- grep(specifiedString, rownames(spectral_data))
#
# # Exclude the matching rows
newData <- spectral_data[-matchingRows, ]
#
# # Output the new data frame
spectral_data <- newData
newData <- NULL
```


Get rid of the x in front of the wavelengths
```{r}
updatedColumns <- colnames(spectral_data)
for (i in seq_along(updatedColumns)) {
  updatedColumns[i] <- substring(updatedColumns[i], 2)
}
colnames(spectral_data) <- updatedColumns
```


To run the hsdar package, we have to turn our csv file into speclib
csv to speclib code adopted from: https://www.kaggle.com/code/kaiesalmahmud/csv-to-speclib-updated/notebook
```{r}
# converting the Data Frame into Matrix
library(hsdar)

                                  # Change spectral_data to your own data frame
spectral_data_matrix <- as.matrix(spectral_data)

colnames(spectral_data_matrix) <- NULL
rownames(spectral_data_matrix) <- NULL

wavelengths = names(spectral_data)
wavelengths = as.numeric(wavelengths)
newSpeclib <- speclib(spectral_data_matrix, wavelengths)
str(newSpeclib)

```


Sample plots
```{r}
par(mfrow = c(2,2))
plot(newSpeclib, FUN = 1, main = "First spectrum of Speclib")
plot(newSpeclib, FUN = "median", main = "Median spectrum")
plot(newSpeclib, FUN = "mean", main = "Mean spectrum")

```

Split data set by site code (e.g LIB, BIG, KAT, SHE, etc.)
```{r}
# Get the first 11 characters of the row names
row_names_prefix <- substr(rownames(spectral_data), 1, 11)

# Use the split function to group rows based on the first 11 characters
all_plants <- split(spectral_data, row_names_prefix)
all_plants # <-- this is a list of data frames // when we call it (e.g all_plants[3]) we have to make sure we convert it to a data frame (to.data.frame) for hsdar
```

```{r}
class(all_plants[3]) # have to change it from class type list to daat frame for HSDAR compatibility
```


List of all of the site names(allows us to access each site)
```{r}
data_frame_names <- names(all_plants)
data_frame_names

```



Sanity check /example of calling a specific site / SELECT YOUR PLANT: Make list site data into a data frame 
```{r}
site_data <- all_plants[data_frame_names[1]]
```


FUNCTION TO FIND 5th, 50th, and 95th quartile: e.g function call: quartiles(data_frame_names[8])
```{r}
quartiles <- function(site_name){
  
      #===== list to data frame
      site_data <- all_plants[site_name]
      quartile_df <- as.data.frame(site_data)
      
      #===== fix names of columns
      # Get the current column names
      current_colnames <- colnames(quartile_df)
      
      # Remove the first 12 characters from the column names
      new_colnames <- substr(current_colnames, 13, nchar(current_colnames))
      
      # Update the column names of the data frame
      colnames(quartile_df) <- new_colnames
      
      #===== Quartiles calculations // output is a data frame
      # Calculate the quartiles for each column // update what you want to calculate here
      quartiles <- apply(quartile_df, 2, function(x) quantile(x, probs = c(0.05, 0.5, 0.95)))
      
      # Create a new data frame to store the results
      result_data_frame <- data.frame(
        "5th_Quartile" = quartiles[1, ],
        "50th_Quartile" = quartiles[2, ],
        "95th_Quartile" = quartiles[3, ]
      )
      
      # Transpose the data frame (flip rows and columns) and convert back to data frame
      quartile_df <- as.data.frame(t(result_data_frame))
      
      #Remove the "X" in the row names
      updatedRows <- row.names(quartile_df)
      for (i in seq_along(updatedRows)) {
        updatedRows[i] <- substring(updatedRows[i], 2)
      }
      row.names(quartile_df) <- updatedRows
      
      quartile_df <- as.data.frame(t(quartile_df))
      
      # Convert row names to a new column and reset the row index
      quartile_df$Wavelength <- rownames(quartile_df)
      
      # Remove the current index column from the dataframe
      rownames(quartile_df) <- NULL
      
      quartile_df <- quartile_df %>%
      select("Wavelength", everything()) %>%
      arrange("Wavelength")
            
      # Print the transposed data frame
      return(quartile_df)
  
}
```


QUARTILE PLOT FUNCTION
```{r}
#== define what quartile you want to plot (1 = 5th, 2 = 50th, 3 = 95th)
 plotSpectra <- function(data, quartile, siteName) {
   
   
  wavelength <- data[[1]]
  sample <- data[, -1]
  
    current_dependent <- sample[[quartile]]
    
    plot_title <- paste(names(sample)[quartile], ": ", siteName)
    plot(x = wavelength, y = current_dependent, 
         main = plot_title, xlab = "Wavelength (nm", ylab = "Reflectance %", type = "l")
  
}
```


function call quartile and plot (sanity check)
```{r}
name <- data_frame_names[2]
plotSpectra(quartiles(data_frame_names[2]),1, name)

```



MEAN function: e.g function call: mean(data_frame_names[8])
```{r}
mean_fn <- function(siteName){
    # List to data frame
    #site_data <- all_plants[data_frame_names[1]]
    site_data <- all_plants[siteName]
    mean_df <- as.data.frame(do.call(cbind, site_data))
    
    # Calculate the mean for each column (wavelength)
    mean_values <- colMeans(mean_df, na.rm = TRUE)
    
    
    # Create a new data frame to store the mean values
    mean_result_df <- data.frame("mean" = mean_values)
    
    # Transpose the data frame (flip column and rows)
    transposed_mean_df <- t(mean_result_df)
    
    # Convert the transposed matrix back to a data frame
    transposed_mean_df <- as.data.frame(transposed_mean_df)
    
    # Get rid of the sample names
    current_colnames <- colnames(transposed_mean_df)
    new_colnames <- substr(current_colnames, 13, nchar(current_colnames))
    colnames(transposed_mean_df) <- new_colnames
    
    # Print the transposed data frame
    return(transposed_mean_df)
}
```


MEDIAN function: e.g function call: median_fn(data_frame_names[8])
```{r}
median_fn <- function(siteName){
    # List to data frame
    site_data <- all_plants[siteName]
    median_df <- as.data.frame(do.call(cbind, site_data))
    
    # Calculate the median for each column (wavelength)
    median_values <- apply(median_df, 2, median, na.rm = TRUE)
    
    # Create a new data frame to store the median values
    median_result_df <- data.frame("median" = median_values)
    
    # Transpose the data frame (flip column and rows)
    transposed_median_df <- t(median_result_df)
    
    # Convert the transposed matrix back to a data frame
    transposed_median_df <- as.data.frame(transposed_median_df)
    
    # Get rid of the sample names
    current_colnames <- colnames(transposed_median_df)
    new_colnames <- substr(current_colnames, 13, nchar(current_colnames))
    colnames(transposed_median_df) <- new_colnames
    
    # Print the transposed data frame
    return(transposed_median_df)
}

```

function call mean and median (sanity check)
```{r}
x <- median_fn(data_frame_names[8])
x
y <- mean_fn(data_frame_names[8])
y
```


FUNCTION (HSDAR): DATA TO SPECLIB
```{r}
## Function parameter: specify the site data frame (e.g all_plants[data_frame_names[i]])
data_to_speclib <- function(site_df) {
  
  if (class(site_df) == "list"){
    site_df <- as.data.frame(site_df)

    # Get the current column names
    current_colnames <- colnames(site_df)
    
    # Remove the first 12 characters from the column names
    new_colnames <- substr(current_colnames, 13, nchar(current_colnames))
    
    # Update the column names of the data frame
    colnames(site_df) <- new_colnames
  }
  
  spectral_data_matrix <- as.matrix(site_df)

  colnames(spectral_data_matrix) <- NULL
  rownames(spectral_data_matrix) <- NULL

  wavelengths = names(site_df)

  wavelengths = as.numeric(wavelengths)
  newSpeclib <- speclib(spectral_data_matrix, wavelengths)

  spectral_data_matrix <- NULL
  wavelengths <- NULL
  
  return(newSpeclib)

}
```

function call: median, mean, standard deviation plots (sanity check)
```{r}
site_data <- all_plants[data_frame_names[1]]

specLib <- data_to_speclib(site_data)

par(mfrow = c(2, 2))
par(pty = "m")

# Plot each plot in its corresponding panel
#plot(specLib, FUN = 1, main = "First spectrum of Speclib")
plot(specLib, FUN = "median", main = "Median spectrum")
plot(specLib, main = "Mean Spectrum with Standard Deviation")

#plot(specLib, FUN = "min", main = "Minimum of speclib")
#plot(specLib, FUN = "max", main = "Maximum of speclib")


```

SPECTRAL COMPARISION FUNCTION
```{r}
start <- NULL
increment <- NULL
step <- NULL

#empty data frame to store spectral comparison values once calculated + the sample site names
result_df <- data.frame(Name = character(), Calculation = numeric(), stringsAsFactors = FALSE)


## type = ii, ip, or pp // interface vs interace, aquatic vs aquatic, aquatic vs interface
# technique = euclidean or sam (spectral angle mapper)
# summary method = mean_fn or median_fn // these are functions previously defined (i.e we're taking the mead/median of the data at each
#                                                                                     wavelength then comparing them to other samples)

comparision_technique <- function(type, technique, summary_method){
  
  if (type == "ii"){
    start <- 1
    increment <- 4
    step <- 1 # <- [i + 2] <- [i + 1 + 1] <- [i + 1 + step]
  }
  else if (type == "pp"){
    start <- 2
    increment <- 4
    step <- 1 # <- [i + 2]
  }
  else if (type == "ip"){
    start <- 1
    increment <- 2
    step <- 0 # <- [i + 1]
  }
  else if (type== "pi"){
    start <- 1
    increment <- 2
    step <- 0 # <- [i + 1]
    }
  else{
    print("invalid type of comparision (e.g ii [interface vs interface])")
  }
  
  for (i in seq(from=start, to=length(data_frame_names), by = increment)){
  
    site1 <- data_frame_names[i]
    site2 <- data_frame_names[i + 1 + step]
    
    site1_summary <- summary_method(site1)
    site2_summary <- summary_method(site2)
    
    site_combined <- rbind(site1_summary, site2_summary)
    
    site_combined <- data_to_speclib(site_combined)
    
    comparison <- dist.speclib(site_combined, method = technique)
    
    sites_compared_name <- paste0(site1,"_", site2)
    
    # Create a new row for each iteration             # change column name here / SAM to Euclidean or Euclidean to SAM
    new_row <- data.frame(Sites = sites_compared_name, Euclidean = comparison)

    # Append the new row to the result_df data frame
    result_df <- rbind(result_df, new_row)
    ## take out
    #print(paste0(site1, " & ", site2, ": ", comparison))
  }
  
  return(result_df)

}

```


Function call
```{r}

spectral_comp_df <- comparision_technique("pp", "euclidean", median_fn)

```

Save results to a csv
```{r}

file_path <- "/Users/charles/Desktop/wp_asd_spectra/wp_spring_2022/Spectral_comparison/spring_pp_median_euclidean.csv"


# Use the write.csv function to save the data frame as a CSV
write.csv(spectral_comp_df, file = file_path, row.names = FALSE)

```


PLOT MEAN AND MEDIAN
```{r}

# data_frame_names is a list of names for all of the sites
for (i in seq_along(data_frame_names)) {
  
  # Accessing the site name within the data_frame_names (a list)
  site_name <- data_frame_names[[i]]
  
  # Make new data frame with site name // all_plants is a list that contains all of the data frames for each site
  site_df <- all_plants[[site_name]]
  
  # Make it into specLib (hsdar)
  specLib <- data_to_speclib(site_df)
  quartile_site <- quartiles(site_name)

  # Plot
  par(mfrow = c(2, 3))
  par(pty = "s")
  #par(pty = "m")

  plot(specLib, FUN = 1, main = paste("First spectrum: ", site_name))
  plotSpectra(quartile_site,1, site_name)
  plot(specLib, FUN = "median", main = paste("Median spectrum:", site_name))
  plotSpectra(quartile_site,2, site_name)
  #plot(specLib, FUN = "mean",main = paste("Mean spectrum:", site_name))
  plot(specLib, main = paste("Mean spectrum:", site_name))
  plotSpectra(quartile_site,3, site_name)
}

```

Save plots above to pdf
```{r}
# Define the file path and name for the PDF file
pdf_file_path <- "/Users/charles/Desktop/wp_asd_spectra/wp_spring_2022/wp_summary_plots_WR_removed.pdf"

# Start the PDF file
pdf(pdf_file_path)

# Number of plots per page
plots_per_page <- 3

# Outer loop to iterate over sets of four sites
for (i in seq(1, length(data_frame_names), plots_per_page)) {
  # Set up the plotting layout for four plots per page
  par(mfrow = c(plots_per_page, 1))

  # Inner loop to generate plots for each site in the set
  for (j in seq(i, min(i + plots_per_page - 1, length(data_frame_names)))) {
    # Accessing the site name within the data_frame_names (a list)
    site_name <- data_frame_names[[j]]
    
    # Make new data frame with site name // all_plants is a list that contains all of the data frames for each site
    site_df <- all_plants[[site_name]]
    
    # Make it into specLib (hsdar)
    specLib <- data_to_speclib(site_df)
    quartile_site <- quartiles(site_name)

    # Plot the three plots in a single row
    plot(specLib, FUN = 1, main = paste("First spectrum: ", site_name))
    plotSpectra(quartile_site,1, site_name)
    plot(specLib, FUN = "median", main = paste("Median spectrum:", site_name))
    plotSpectra(quartile_site,2, site_name)
    #plot(specLib, FUN = "mean",main = paste("Mean spectrum:", site_name))
    plot(specLib, main = paste("Mean spectrum:", site_name))
    plotSpectra(quartile_site,3, site_name)
  }
}

# End the PDF file
dev.off()

```


Splicing data and taking out bad spectra

```{r}
#========== Make index

# Add row numbers as a new column
spectral_data$Row_Index <- seq_len(nrow(spectral_data))

# Convert row names to a new column and reset the row index
spectral_data$SampleName <- rownames(spectral_data)

# Remove the current index column from the dataframe
rownames(spectral_data) <- NULL
# Reorder the columns to have SampleName as the first column
spectral_data <- spectral_data[, c("Row_Index", "SampleName", names(spectral_data)[-c(2, ncol(spectral_data))])]


#============
output_file_path <- "/Users/charles/Desktop/wp_asd_spectra/wp_summer_2021/output_file.csv"

#spectral_data <- spectral_data[, -which(names(spectral_data) == "label")]
spectral_data <- spectral_data[, -which(names(spectral_data) == "Row_Index.1")]


#========== Exclude bad spectra

# Replace 'spectral_data' with your actual dataframe name
# Exclude rows with the specified indices (inclusive to end points)
indices_to_exclude <- c(463:465)
spectral_data_excluded <- spectral_data %>%
  slice(-indices_to_exclude)


```

```{r}
#==========

# Replace 'spectral_data_excluded' with your actual dataframe name
# Replace 'SampleName' with the column name that you want to set as the new index column

spectral_data_excluded <- spectral_data_excluded[,-1]  # Remove the first column (current index column)

rownames(spectral_data_excluded) <- spectral_data_excluded$SampleName  # Set 'SampleName' as the new index
spectral_data <- spectral_data_excluded[, -which(names(spectral_data_excluded) == "SampleName")]
spectral_data
```



